---
title: "TFG CODE for 50 assets"
author: "Guillem Barniol i Albert Sala"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### IMPORTING DATA

The following libraries are needed to run the entire code; however using data provided in the data files allows you to save, probably, some hours of computations.

```{r}
library(readxl)
library(quantmod)
library(nlshrink)
library(RiskPortfolios)
suppressPackageStartupMessages(library(xts))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(matrixStats))
```

In the following file we have the constituents of S&P500 at date ....

```{r}
#data <- read_excel("sp500.xlsx")
#attach(data)
```

Using library *quantmod* we import data from Yahoo for the S&P500 constituents. We drop all stocks with at least one NA value and then we compute the log return for each stock. To reduce computational time one can import directly the data file "X" where the log returns of all final assets considered are saved.

```{r}
#tickers <- c(Symbol)
#starting <- "2000-01-01"
#ending <- "2021-02-27"
#stockPrices <- NULL

#for (ticker in tickers)
#        stockPrices <- cbind(stockPrices, getSymbols(ticker, from=starting, to=ending, auto.assign=FALSE)[,6])

#stockPrices <- t(na.omit(t(stockPrices)))
#X <- diff(log(stockPrices))*100

#write.table(X,'../TFG/companies_log_returns.csv', sep=' ', row.names=TRUE, col.names=TRUE)
X <- as.matrix(read.table('../TFG/companies_log_returns.csv', header=TRUE, sep=' '))
```

We download the S&P500 index prices to compute log returns; they are defined as the market factor returns. Again one can import data from *excel* which is faster.

```{r}
#sp500 <- cbind(getSymbols("^GSPC", from=starting, to=ending, auto.assign = FALSE)[,6])
#sp500 <- (diff(log(sp500))*100)[-1]

#write.table(sp500,'../TFG/S&P500_log_returns.csv', sep=' ', row.names=TRUE, col.names=TRUE)
sp500 <- as.matrix(read.table('../TFG/S&P500index_log_returns.csv', header=TRUE, sep=' '))
```

Also, we need to import Fama-French factors. Here we download an *excel* data file from Kenneth R. French data library and select the period we are interested in. Note we substitute the risk premium for the S&P500 index returns, which we define as the market.

```{r}
fama_factors <- as.matrix(read.table('../TFG/5 factors.csv', header=TRUE, sep=','))
rownames(fama_factors) <- fama_factors[,1]
fama_factors <- fama_factors[,2:6]
fama_factors[,1] <- sp500
```

Here one can choose to perform the analysis with all 378 assets or with a random selection of less of the 378 assets. Remember **always** to run previously to the random selection the *set.seed* command if you want to get the exact same results that we report.

```{r}
random <- 50   #choose the number of assets you want to take only if you don't want to run the computations for all the assets
set.seed(1312)
X <- X[, sample(ncol(X), random)]
```

### CONSTRUCTING GMV portfolio given the covariance estimator

The bellow computations makes easier to write the rest of the code.

```{r}
T <- dim(X)[1] #number of trading days
N <- dim(X)[2] #number of assets
a <- 1004      #(arbitrary) number of days we look back to compute covariance matrix estimator
ones <-matrix(rep(1,N),N,1) #the appropiate vector of ones to compute GMV portfolio weights
dates <- as.Date(rownames(X))
```

The following code imports directly the returns and asset's weights for each portfolio and is much faster than computing them through the original code that follows afterwards.

```{r}
#1. EQUAL WEIGHT
demiguelR <- as.matrix(read.table('../TFG/Resultats/50/DeMiguel R.csv', header=TRUE, sep=' '))
demiguelW <- as.matrix(read.table('../TFG/Resultats/50/DeMiguel W.csv', header=TRUE, sep=' '))

#2. SAMPLE COVARIANCE
sampleR <- as.matrix(read.table('../TFG/Resultats/50/Sample_cov R.csv', header=TRUE, sep=' '))
sampleW <- as.matrix(read.table('../TFG/Resultats/50/Sample_cov W.csv', header=TRUE, sep=' '))

#3. LINEAR SHRINKAGE to IDENTITY MATRIX
wolfR <- as.matrix(read.table('../TFG/Resultats/50/LR sam_cov R.csv', header=TRUE, sep=' '))
wolfW <- as.matrix(read.table('../TFG/Resultats/50/LR sam_cov W.csv', header=TRUE, sep=' '))

#4. NON-LINEAR SHRINKAGE to IDENTITY MATRIX
sampleNlR <- as.matrix(read.table('../TFG/Resultats/50/NL sam_cov R.csv', header=TRUE, sep=' '))
sampleNlW <- as.matrix(read.table('../TFG/Resultats/50/NL sam_cov W.csv', header=TRUE, sep=' '))

#6. MARKET FACTOR with LINEAR SHRINKAGE IN RESIDUALS
marketR <- as.matrix(read.table('../TFG/Resultats/50/Market LR residuals R.csv', header=TRUE, sep=' '))
marketW <- as.matrix(read.table('../TFG/Resultats/50/Market LR residuals W.csv', header=TRUE, sep=' '))

#7. FAMA FRENCH 3 FACTORS with LINEAR SHRINKAGE IN RESIDUALS
fama3R <- as.matrix(read.table('../TFG/Resultats/50/Fama3 LR residuals R.csv', header=TRUE, sep=' '))
fama3W <- as.matrix(read.table('../TFG/Resultats/50/Fama3 LR residuals W.csv', header=TRUE, sep=' '))

#8.  FAMA FRENCH 5 FACTORS with LINEAR SHRINKAGE IN RESIDUALS
fama5R <- as.matrix(read.table('../TFG/Resultats/50/Fama5 LR residuals R.csv', header=TRUE, sep=' '))
fama5W <- as.matrix(read.table('../TFG/Resultats/50/Fama5 LR residuals W.csv', header=TRUE, sep=' '))

#9. LINEAR SHRINKAGE to CONSTANT CORRELATION (LEDOIT and WOLF)
con_wolfR <- as.matrix(read.table('../TFG/Resultats/50/Con_corr LW R.csv', header=TRUE, sep=' '))
con_wolfW <- as.matrix(read.table('../TFG/Resultats/50/Con_corr LW W.csv', header=TRUE, sep=' '))

#11. PCA with 1 PCs
pca1R <- as.matrix(read.table('../TFG/Resultats/50/PCA1 LR residuals R.csv', header=TRUE, sep=' '))
pca1W <- as.matrix(read.table('../TFG/Resultats/50/PCA1 LR residuals W.csv', header=TRUE, sep=' '))

#12. PCA with 3 PCs
pca3R <- as.matrix(read.table('../TFG/Resultats/50/PCA3 LR residuals R.csv', header=TRUE, sep=' '))
pca3W <- as.matrix(read.table('../TFG/Resultats/50/PCA3 LR residuals W.csv', header=TRUE, sep=' '))

#13. PCA with 5 PCs
pca5R <- as.matrix(read.table('../TFG/Resultats/50/PCA5 LR residuals R.csv', header=TRUE, sep=' '))
pca5W <- as.matrix(read.table('../TFG/Resultats/50/PCA5 LR residuals W.csv', header=TRUE, sep=' '))
```

1. EQUAL WEIGHT

```{r}
demiguelWi <- NULL
demiguelW <- NULL
demiguelRi <-NULL
demiguelR <-NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        demiguelWi <- matrix(rep(1/N), N, 1)
        demiguelW<-cbind(demiguelW,demiguelWi)}
        else{
        demiguelWi <- demiguelW[,t-a-1]*(1+X[t-1,]/100)
        demiguelWi <- demiguelWi/sum(demiguelWi)
        demiguelW <- cbind(demiguelW, demiguelWi)}
}

for (t in (a+1):T){
        demiguelRi <- X[t,]%*%demiguelW[,(t-a)]
        demiguelR <-cbind(demiguelR, demiguelRi)
}
demiguelW <- t(demiguelW)
rownames(demiguelW)<-dates[(a+1):T]
demiguelR<-t(demiguelR)
rownames(demiguelR)<-dates[(a+1):T]
colnames(demiguelR)<-"Equal weight"

write.table(demiguelW,'../TFG/Resultats/50/DeMiguel W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(demiguelR,'../TFG/Resultats/50/DeMiguel R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

2. SAMPLE COVARIANCE

```{r}
sampleWi <- NULL
sampleW <- NULL
sampleRi <-NULL
sampleR <-NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){ 
        S_t <- cov(X[(t-a):(t-1),]) 
        sampleWi<-solve(S_t)%*%ones/as.numeric(t(ones)%*%solve(S_t)%*%ones) 
        sampleW<-cbind(sampleW,sampleWi)}  
        else{
        sampleWi <- sampleW[,t-a-1]*(1+X[t-1,]/100) 
        sampleWi <- sampleWi/sum(sampleWi) 
        sampleW <- cbind(sampleW, sampleWi)}  
}

for (t in (a+1):T){
        sampleRi <- X[t,]%*%sampleW[,(t-a)]
        sampleR <-cbind(sampleR, sampleRi)
}
sampleW <- t(sampleW)
rownames(sampleW)<-dates[(a+1):T]
sampleR<-t(sampleR)
rownames(sampleR)<-dates[(a+1):T]
colnames(sampleR)<-"Sample covariance"

write.table(sampleW,'../TFG/Resultats/50/Sample_cov W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(sampleR,'../TFG/Resultats/50/Sample_cov R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

3. LINEAR SHRINKAGE

```{r}
wolfWi <- NULL
wolfW <- NULL
wolfRi <-NULL
wolfR <-NULL
# create a data matrix derived from the original data matrix X with 1004 observations.
for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        XX <- X[(t-a):(t-1),]
        wolfS <- linshrink_cov(XX)
        wolfWi<-solve(wolfS)%*%ones/as.numeric(t(ones)%*%solve(wolfS)%*%ones)
        wolfW<-cbind(wolfW, wolfWi)}
        else{
        wolfWi <- wolfW[,t-a-1]*(1+X[t-1,]/100)
        wolfWi <- wolfWi/sum(wolfWi)
        wolfW <- cbind(wolfW, wolfWi)}
}   

for (t in (a+1):T){
        wolfRi <- X[t,]%*%wolfW[,(t-a)]
        wolfR <-cbind(wolfR, wolfRi)
}

wolfW <- t(wolfW)
rownames(wolfW)<-dates[(a+1):T]
wolfR<-t(wolfR)
rownames(wolfR)<-dates[(a+1):T]
colnames(wolfR)<-"Linear shrinkage"

write.table(wolfW,'../TFG/Resultats/50/LR sam_cov W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(wolfR,'../TFG/Resultats/50/LR sam_cov R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

4.NON-LINEAR SHRINKAGE

```{r}
sampleNLWi <- NULL
sampleNlW <- NULL
sampleNlRi <-NULL
sampleNlR <-NULL
# create a data matrix derived from the original data matrix X with 1004 observations.
for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        wolfS <- nlshrink_cov(X[(t-a):(t-1),])
        sampleNLWi<-solve(wolfS)%*%ones/as.numeric(t(ones)%*%solve(wolfS)%*%ones)
        sampleNlW<-cbind(sampleNlW, sampleNLWi)}
        else{
        sampleNLWi <- sampleNlW[,t-a-1]*(1+X[t-1,]/100)
        sampleNLWi <- sampleNLWi/sum(sampleNLWi)
        sampleNlW <- cbind(sampleNlW, sampleNLWi)}
}   

for (t in (a+1):T){
        sampleNlRi <- X[t,]%*%sampleNlW[,(t-a)]
        sampleNlR <-cbind(sampleNlR, sampleNlRi)
}

sampleNlW <- t(sampleNlW)
rownames(sampleNlW)<-dates[(a+1):T]
sampleNlR<-t(sampleNlR)
rownames(sampleNlR)<-dates[(a+1):T]
colnames(sampleNlR)<-"Non-linear shrinkage"

write.table(sampleNlW,'../TFG/Resultats/50/NL sam_cov W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(sampleNlR,'../TFG/Resultats/50/NL sam_cov R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

5. Market factor with DIAGONALIZING RESIDUALS

```{r}
diag1fWi <- NULL
diag1fW <- NULL
diag1fRi <- NULL
diag1fR <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                capm <- lm(X[((t-a)):(t-1),] ~ sp500[(t-a):(t-1),])
                cov_market <- capm$coefficients[2,]%*%as.matrix(var(sp500[(t-a):(t-1),]))%*%t(capm$coefficients[2,])
                resi <- capm$residuals
                cov_res <- diag(mean(diag(resi)), ncol=ncol(resi), nrow=ncol(resi))  # residuals covariance shrinked
                cov_est <- cov_market + cov_res
                
                diag1fWi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                diag1fW <- cbind(diag1fW, diag1fWi)
        }
        
        else{
                diag1fWi <- diag1fW[,t-a-1]*(1+X[t-1,]/100)
                diag1fWi <- diag1fWi/sum(diag1fWi)
                diag1fW <- cbind(diag1fW, diag1fWi)
        }
}

for (t in (a+1):T){
        diag1fRi <- X[t,]%*%diag1fW[,(t-a)]
        diag1fR <-cbind(diag1fR, diag1fRi)
}

diag1fW <- t(diag1fW)
rownames(diag1fW)<-dates[(a+1):T]
diag1fR<-t(diag1fR)
rownames(diag1fR)<-dates[(a+1):T]
colnames(diag1fR)<-"Market f. and residuals diagonalized"
```

6. MARKET FACTOR with LINEAR SHRINKAGE IN RESIDUALS

```{r}
marketWi <- NULL
marketW <- NULL
marketRi <- NULL
marketR <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                capm <- lm(X[((t-a)):(t-1),] ~ sp500[(t-a):(t-1)])
                cov_market <- capm$coefficients[2,]%*%as.matrix(var(sp500[(t-a):(t-1),1]))%*%t(capm$coefficients[2,])
                cov_res <- linshrink_cov(capm$residuals)  # residuals covariance shrinked
                cov_est <- cov_market + cov_res
                
                marketWi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                marketW <- cbind(marketW, marketWi)
        }
        
        else{
                marketWi <- marketW[,t-a-1]*(1+X[t-1,]/100)
                marketWi <- marketWi/sum(marketWi)
                marketW <- cbind(marketW, marketWi)
        }
}

for (t in (a+1):T){
        marketRi <- X[t,]%*%marketW[,(t-a)]
        marketR <-cbind(marketR, marketRi)
}

marketW <- t(marketW)
rownames(marketW)<-dates[(a+1):T]
marketR<-t(marketR)
rownames(marketR)<-dates[(a+1):T]
colnames(marketR)<-"Market factor + LS"

write.table(marketW,'../TFG/Resultats/50/Market LR residuals W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(marketR,'../TFG/Resultats/50/Market LR residuals R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

7. FAMA-FRENCH three factors with residuals linear shrinkage.

```{r}
fama3Wi <- NULL
fama3W <- NULL
fama3Ri <- NULL
fama3R <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                capm <- lm(X[((t-a)):(t-1),] ~ fama_factors[(t-a):(t-1),1:3])
                cov_market <- t(capm$coefficients[2:4,])%*%cov(fama_factors[(t-a):(t-1),1:3])%*%capm$coefficients[2:4,]
                cov_res <- linshrink_cov(capm$residuals)  # residuals covariance shrinked
                cov_est <- cov_market + cov_res
                
                fama3Wi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                fama3W <- cbind(fama3W, fama3Wi)
        }
        
        else{
                fama3Wi <- fama3W[,t-a-1]*(1+X[t-1,]/100)
                fama3Wi <- fama3Wi/sum(fama3Wi)
                fama3W <- cbind(fama3W, fama3Wi)
        }
}

for (t in (a+1):T){
        fama3Ri <- X[t,]%*%fama3W[,(t-a)]
        fama3R <-cbind(fama3R, fama3Ri)
}

fama3W <- t(fama3W)
rownames(fama3W)<-dates[(a+1):T]
fama3R<-t(fama3R)
rownames(fama3R)<-dates[(a+1):T]
colnames(fama3R)<-"FF3 + LS"

write.table(fama3W,'../TFG/Resultats/50/Fama3 LR residuals W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(fama3R,'../TFG/Resultats/50/Fama3 LR residuals R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

8. FAMA-FRENCH five factors with residuals linear shrinkage.

```{r}
fama5Wi <- NULL
fama5W <- NULL
fama5Ri <- NULL
fama5R <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                capm <- lm(X[((t-a)):(t-1),] ~ fama_factors[(t-a):(t-1),])
                cov_market <- t(capm$coefficients[2:6,])%*%cov(fama_factors[(t-a):(t-1),])%*%capm$coefficients[2:6,]
                cov_res <- linshrink_cov(capm$residuals)  # residuals covariance shrinked
                cov_est <- cov_market + cov_res
                
                fama5Wi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                fama5W <- cbind(fama5W, fama5Wi)
        }
        
        else{
                fama5Wi <- fama5W[,t-a-1]*(1+X[t-1,]/100)
                fama5Wi <- fama5Wi/sum(fama5Wi)
                fama5W <- cbind(fama5W, fama5Wi)
        }
}

for (t in (a+1):T){
        fama5Ri <- X[t,]%*%fama5W[,(t-a)]
        fama5R <-cbind(fama5R, fama5Ri)
}

fama5W <- t(fama5W)
rownames(fama5W)<-dates[(a+1):T]
fama5R<-t(fama5R)
rownames(fama5R)<-dates[(a+1):T]
colnames(fama5R)<-"FF5 + LS"

write.table(fama5W,'../TFG/Resultats/50/Fama5 LR residuals W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(fama5R,'../TFG/Resultats/50/Fama5 LR residuals R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

9. LINEAR SHRINKAGE to CONSTANT CORRELATION (LEDOIT and WOLF)

```{r}
con_wolfWi <- NULL
con_wolfW <- NULL
con_wolfRi <-NULL
con_wolfR <-NULL

# create a data matrix derived from the original data matrix X with 1004 observations.
for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        wolfS <- covEstimation(X[(t-a):(t-1),], control=list(type='cor'))
        con_wolfWi<-solve(wolfS)%*%ones/as.numeric(t(ones)%*%solve(wolfS)%*%ones)
        con_wolfW<-cbind(con_wolfW, con_wolfWi)}
        else{
        con_wolfWi <- con_wolfW[,t-a-1]*(1+X[t-1,]/100)
        con_wolfWi <- con_wolfWi/sum(con_wolfWi)
        con_wolfW <- cbind(con_wolfW, con_wolfWi)}
}   

for (t in (a+1):T){
        con_wolfRi <- X[t,]%*%con_wolfW[,(t-a)]
        con_wolfR <-cbind(con_wolfR, con_wolfRi)
}

con_wolfW <- t(con_wolfW)
rownames(con_wolfW)<-dates[(a+1):T]
con_wolfR<-t(con_wolfR)
rownames(con_wolfR)<-dates[(a+1):T]
colnames(con_wolfR)<-"Constant corr."

write.table(con_wolfW,'../TFG/Resultats/50/Con_corr LW W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(con_wolfR,'../TFG/Resultats/50/Con_corr LW R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

11. PCA with 1 Principal Components

```{r}
pca1Wi <- NULL
pca1W <- NULL
pca1Ri <- NULL
pca1R <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                v <- eigen(cov(X[((t-a)):(t-1),]), symmetric=T)$vectors[,1]
                PCs <- X[((t-a)):(t-1),] %*% v
                capm <- lm(X[((t-a)):(t-1),] ~ PCs)
                cov_pca <- capm$coefficients[2,]%*%cov(PCs)%*%t(capm$coefficients[2,])
                
                cov_res <- linshrink_cov(capm$residuals)  # residuals covariance shrinked
                cov_est <- cov_pca + cov_res
                
                pca1Wi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                pca1W <- cbind(pca1W, pca1Wi)
        }
        
        else{
                pca1Wi <- pca1W[,t-a-1]*(1+X[t-1,]/100)
                pca1Wi <- pca1Wi/sum(pca1Wi)
                pca1W <- cbind(pca1W, pca1Wi)
        }
}

for (t in (a+1):T){
        pca1Ri <- X[t,]%*%pca1W[,(t-a)]
        pca1R <-cbind(pca1R, pca1Ri)
}

pca1W <- t(pca1W)
rownames(pca1W)<-dates[(a+1):T]
pca1R<-t(pca1R)
rownames(pca1R)<-dates[(a+1):T]
colnames(pca1R)<-"1 PC + LS"

write.table(pca1W,'../TFG/Resultats/50/PCA1 LR residuals W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(pca1R,'../TFG/Resultats/50/PCA1 LR residuals R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

12. PCA with 3 PCs

```{r}
k <- 3

pca3Wi <- NULL
pca3W <- NULL
pca3Ri <- NULL
pca3R <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                v <- eigen(cov(X[((t-a)):(t-1),]), symmetric=T)$vectors[,1:k]
                PCs <- X[((t-a)):(t-1),] %*% v
                capm <- lm(X[((t-a)):(t-1),] ~ PCs)
                cov_pca <- t(capm$coefficients[2:(k+1),])%*%cov(PCs)%*%capm$coefficients[2:(k+1),]
                
                cov_res <- linshrink_cov(capm$residuals)  # residuals covariance shrinked
                cov_est <- cov_pca + cov_res
                
                pca3Wi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                pca3W <- cbind(pca3W, pca3Wi)
        }
        
        else{
                pca3Wi <- pca3W[,t-a-1]*(1+X[t-1,]/100)
                pca3Wi <- pca3Wi/sum(pca3Wi)
                pca3W <- cbind(pca3W, pca3Wi)
        }
}

for (t in (a+1):T){
        pca3Ri <- X[t,]%*%pca3W[,(t-a)]
        pca3R <-cbind(pca3R, pca3Ri)
}

pca3W <- t(pca3W)
rownames(pca3W)<-dates[(a+1):T]
pca3R<-t(pca3R)
rownames(pca3R)<-dates[(a+1):T]
colnames(pca3R)<-"3 PCs + LS"

write.table(pca3W,'../TFG/Resultats/50/PCA3 LR residuals W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(pca3R,'../TFG/Resultats/50/PCA3 LR residuals R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

13. PCA with 5 PCs

```{r}
k <- 5

pca5Wi <- NULL
pca5W <- NULL
pca5Ri <- NULL
pca5R <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
                v <- eigen(cov(X[((t-a)):(t-1),]), symmetric=T)$vectors[,1:k]
                PCs <- X[((t-a)):(t-1),] %*% v
                capm <- lm(X[((t-a)):(t-1),] ~ PCs)
                cov_pca <- t(capm$coefficients[2:(k+1),])%*%cov(PCs)%*%capm$coefficients[2:(k+1),]
                
                cov_res <- linshrink_cov(capm$residuals)  # residuals covariance shrinked
                cov_est <- cov_pca + cov_res
                
                pca5Wi<-solve(cov_est)%*%ones/as.numeric(t(ones)%*%solve(cov_est)%*%ones)
                pca5W <- cbind(pca5W, pca5Wi)
        }
        
        else{
                pca5Wi <- pca5W[,t-a-1]*(1+X[t-1,]/100)
                pca5Wi <- pca5Wi/sum(pca5Wi)
                pca5W <- cbind(pca5W, pca5Wi)
        }
}

for (t in (a+1):T){
        pca5Ri <- X[t,]%*%pca5W[,(t-a)]
        pca5R <-cbind(pca5R, pca5Ri)
}

pca5W <- t(pca5W)
rownames(pca5W)<-dates[(a+1):T]
pca5R<-t(pca5R)
rownames(pca5R)<-dates[(a+1):T]
colnames(pca5R)<-"5 PCs + LS"

write.table(pca5W,'../TFG/Resultats/50/PCA5 LR residuals W.csv', sep=' ', row.names=TRUE, col.names=TRUE)
write.table(pca5R,'../TFG/Resultats/50/PCA5 LR residuals R.csv', sep=' ', row.names=TRUE, col.names=TRUE)
```

### RESULTS

Average return and volatility per year for the portfolios. We compute IR as return/volatility as in Ledoit and Wolf (2018).

```{r eval=FALSE, include=FALSE}
#EQUAL WEIGHT
demiguelR_mean <-((cumprod(1+demiguelR/100)[T-a]^(1/(T-a)))^252-1)*100
demiguelR_vol <- (sd(demiguelR/100))*sqrt(252)*100
demiguelIR <- demiguelR_mean/demiguelR_vol

#SAMPLE COVARIANCE
sampleR_mean <-((cumprod(1+sampleR/100)[T-a]^(1/(T-a)))^252-1)*100
sampleR_vol <- (sd(sampleR/100))*sqrt(252)*100
sampleIR <- sampleR_mean/sampleR_vol

#LINEAR SHRINKAGE
wolfR_mean <-((cumprod(1+wolfR/100)[T-a]^(1/(T-a)))^252-1)*100
wolfR_vol <- (sd(wolfR/100))*sqrt(252)*100
wolfIR <- wolfR_mean/wolfR_vol

#Non-linear
sampleNlR_mean <-((cumprod(1+sampleNlR/100)[T-a]^(1/(T-a)))^252-1)*100
sampleNlR_vol <- (sd(sampleNlR/100))*sqrt(252)*100
sampleNlIR <- sampleNlR_mean/sampleNlR_vol

#MARKET FACTOR + RESIDUALS SHRINKAGE
marketR_mean <-((cumprod(1+marketR/100)[T-a]^(1/(T-a)))^252-1)*100
marketR_vol <- (sd(marketR/100))*sqrt(252)*100
IRmarket <- marketR_mean/marketR_vol

#FAMA-FRENCH (3) FACTORS + RESIDUALS SHRINKAGE
fama3R_mean <-((cumprod(1+fama3R/100)[T-a]^(1/(T-a)))^252-1)*100
fama3R_vol <- (sd(fama3R/100))*sqrt(252)*100
IR3fama <- fama3R_mean/fama3R_vol

#FAMA-FRENCH (5) FACTORS + RESIDUALS SHRINKAGE
fama5R_mean <- ((cumprod(1+fama5R/100)[T-a]^(1/(T-a)))^252-1)*100
fama5R_vol <- (sd(fama5R/100))*sqrt(252)*100
fama5IR <- fama5R_mean/fama5R_vol

#SAMPLE COV TO CONSTANT CORRELATION LW2003
con_wolfR_mean <- ((cumprod(1+con_wolfR/100)[T-a]^(1/(T-a)))^252-1)*100
con_wolfR_vol <- (sd(con_wolfR/100))*sqrt(252)*100
con_wolfR_IR <- ((cumprod(1+con_wolfR/100)[T-a]^(1/(T-a)))^252-1)*100/((sd(con_wolfR/100))*sqrt(252)*100)

#PCA1
pca1R_mean <-((cumprod(1+pca1R/100)[T-a]^(1/(T-a)))^252-1)*100
pca1R_vol <- (sd(pca1R/100))*sqrt(252)*100
pca1IR <- pca1R_mean/pca1R_vol

#PCA3
pca3R_mean <-((cumprod(1+pca3R/100)[T-a]^(1/(T-a)))^252-1)*100
pca3R_vol <- (sd(pca3R/100))*sqrt(252)*100
pca3IR <- pca3R_mean/pca3R_vol

#PCA5
pca5R_mean <-((cumprod(1+pca5R/100)[T-a]^(1/(T-a)))^252-1)*100
pca5R_vol <- (sd(pca5R/100))*sqrt(252)*100
pca5IR <- pca5R_mean/pca5R_vol

```

AV (average return per year in %); SD (average volatility per year in %); IR

```{r eval=FALSE, include=FALSE}
AV <- c(demiguelR_mean, sampleR_mean, wolfR_mean, sampleNlR_mean, marketR_mean, fama3R_mean, fama5R_mean, con_wolfR_mean, pca1R_mean, pca3R_mean, pca5R_mean)
SD <- c(demiguelR_vol, sampleR_vol, wolfR_vol, sampleNlR_vol, marketR_vol, fama3R_vol, fama5R_vol, con_wolfR_vol, pca1R_vol, pca3R_vol, pca5R_vol)
IR <- c(demiguelIR, sampleIR, wolfIR, sampleNlIR, IRmarket, IR3fama, fama5IR, con_wolfR_IR, pca1IR, pca3IR, pca5IR)
```

Average (monthly) turnover (TO)

```{r eval=FALSE, include=FALSE}
#EQUAL WEIGHT
demiguel_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(demiguelW[t-a,]-demiguelW[t-a-1,]))
        demiguel_mon_to<-cbind(demiguel_mon_to, suma)}
}
demiguelTO <- sum(demiguel_mon_to)/(floor((T-a)/21)-1)


#SAMPLE COVARIANCE
suma <- NULL
sample_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(sampleW[t-a,]-sampleW[t-a-1,]))
        sample_mon_to<-cbind(sample_mon_to, suma)}
}
sampleTO <- sum(sample_mon_to)/((floor((T-a))/21)-1)

#LINEAR SHRINKAGE
wolf_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(wolfW[t-a,]-wolfW[t-a-1,]))
        wolf_mon_to<-cbind(wolf_mon_to, suma)}
}
wolfTO <- sum(wolf_mon_to)/(floor((T-a)/21)-1)

#Non-linear shrinkage
NL_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(sampleNlW[t-a,]-sampleNlW[t-a-1,]))
        NL_mon_to<-cbind(NL_mon_to, suma)}
}
NLinearTO <- sum(NL_mon_to)/(floor((T-a)/21)-1)

#MARKET FACTOR + LINEAR SHRINKAGE IN RESIDUALS
market_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(marketW[t-a,]-marketW[t-a-1,]))
        market_mon_to<-cbind(market_mon_to, suma)}
}
marketTO <- sum(market_mon_to)/(floor((T-a)/21)-1)

#FAMA FRENCH 3 + LINEAR SHRINKAGE IN RESIDUALS
fama3_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(fama3W[t-a,]-fama3W[t-a-1,]))
        fama3_mon_to<-cbind(fama3_mon_to, suma)}
}
fama3TO <- sum(fama3_mon_to)/(floor((T-a)/21)-1)

#FAMA FRENCH 5 + LINEAR SHRINKAGE IN RESIDUALS
fama5_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(fama5W[t-a,]-fama5W[t-a-1,]))
        fama5_mon_to<-cbind(fama5_mon_to, suma)}
}
fama5TO <- sum(fama5_mon_to)/(floor((T-a)/21)-1)

#Constant correlation + LINEAR SHRINKAGE IN RESIDUALS
conscor_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(con_wolfW[t-a,]-con_wolfW[t-a-1,]))
        conscor_mon_to<-cbind(conscor_mon_to, suma)}
}
conscorTO <- sum(conscor_mon_to)/(floor((T-a)/21)-1)

#PCA 1 + LINEAR SHRINKAGE IN RESIDUALS
pca1_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(pca1W[t-a,]-pca1W[t-a-1,]))
        pca1_mon_to<-cbind(pca1_mon_to, suma)}
}
pca1TO <- sum(pca1_mon_to)/(floor((T-a)/21)-1)

#PCA 3 + LINEAR SHRINKAGE IN RESIDUALS
pca3_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(pca3W[t-a,]-pca3W[t-a-1,]))
        pca3_mon_to<-cbind(pca3_mon_to, suma)}
}
pca3TO <- sum(pca3_mon_to)/(floor((T-a)/21)-1)

#PCA 5 + LINEAR SHRINKAGE IN RESIDUALS
pca5_mon_to <- NULL
for (t in (a+1):T){
        if (!(t-a-1)%%21 & t!=a+1){
        suma <- sum(abs(pca5W[t-a,]-pca5W[t-a-1,]))
        pca5_mon_to<-cbind(pca5_mon_to, suma)}
}
pca5TO <- sum(pca5_mon_to)/(floor((T-a)/21)-1)

TO <- c(demiguelTO, sampleTO, wolfTO, NLinearTO, marketTO, fama3TO, fama5TO, conscorTO, pca1TO, pca3TO, pca5TO)
```

Average (monthly) gross Leverage (GL)

```{r eval=FALSE, include=FALSE}
#EQUAL WEIGHT
demiguel_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(demiguelW[t-a,]))
        demiguel_mon_gl<-cbind(demiguel_mon_gl, suma)}
}
demiguelGL <- sum(demiguel_mon_gl)/(floor((T-a)/21))

#SAMPLE COVARIANCE
sample_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(sampleW[t-a,]))
        sample_mon_gl<-cbind(sample_mon_gl, suma)}
}
sampleGL <- sum(sample_mon_gl)/(floor((T-a))/21)

#LINEAR SHRINKAGE
wolf_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(wolfW[t-a,]))
        wolf_mon_gl<-cbind(wolf_mon_gl, suma)}
}
wolfGL <- sum(wolf_mon_gl)/(floor((T-a))/21)

#Non-linear shrinkage
NL_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(sampleNlW[t-a,]))
        NL_mon_gl<-cbind(NL_mon_gl, suma)}
}
NLinearGL <- sum(NL_mon_gl)/(floor((T-a))/21)

#MARKET FACTOR + LINEAR SHRINKAGE IN RESIDUALS
market_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(marketW[t-a,]))
        market_mon_gl<-cbind(market_mon_gl, suma)}
}
marketGL <- sum(market_mon_gl)/(floor((T-a))/21)

#FAMA FRENCH 3 + LINEAR SHRINKAGE IN RESIDUALS
fama3_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(fama3W[t-a,]))
        fama3_mon_gl<-cbind(fama3_mon_gl, suma)}
}
fama3GL <- sum(fama3_mon_gl)/(floor((T-a))/21)

#FAMA FRENCH 5 + LINEAR SHRINKAGE IN RESIDUALS
fama5_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(fama5W[t-a,]))
        fama5_mon_gl<-cbind(fama5_mon_gl, suma)}
}
fama5GL <- sum(fama5_mon_gl)/(floor((T-a))/21)

#Constant correlation + LINEAR SHRINKAGE IN RESIDUALS
conscor_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(con_wolfW[t-a,]))
        conscor_mon_gl<-cbind(conscor_mon_gl, suma)}
}
conscorGL <- sum(conscor_mon_gl)/(floor((T-a))/21)

#PCA 1 + LINEAR SHRINKAGE IN RESIDUALS
pca1_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(pca1W[t-a,]))
        pca1_mon_gl<-cbind(pca1_mon_gl, suma)}
}
pca1GL <- sum(pca1_mon_gl)/(floor((T-a))/21)

#PCA 3 + LINEAR SHRINKAGE IN RESIDUALS
pca3_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(pca3W[t-a,]))
        pca3_mon_gl<-cbind(pca3_mon_gl, suma)}
}
pca3GL <- sum(pca3_mon_gl)/(floor((T-a))/21)

#PCA 5 + LINEAR SHRINKAGE IN RESIDUALS
pca5_mon_gl <- NULL

for (t in (a+1):T){
        if (t==a+1 | !(t-a-1)%%21){
        suma <- sum(abs(pca5W[t-a,]))
        pca5_mon_gl<-cbind(pca5_mon_gl, suma)}
}
pca5GL <- sum(pca5_mon_gl)/(floor((T-a))/21)



GL <- c(demiguelGL, sampleGL, wolfGL, NLinearGL, marketGL, fama3GL, fama5GL, conscorGL, pca1GL, pca3GL, pca5GL)
```

Average (monthly) proportion of leverage (PL).

```{r}
#EQUAL WEIGHT
demiguel_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(demiguelW[t-a,]<0, 1, 0))/N
    demiguel_mon_pl <- cbind(demiguel_mon_pl, suma)
  }
}
demiguelPL <- sum(demiguel_mon_pl)/(floor((T-a)/21))

#SAMPLE COVARIANCE
sample_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(sampleW[t-a,]<0, 1, 0))/N
    sample_mon_pl <- cbind(sample_mon_pl, suma)
  }
}
samplePL <- sum(sample_mon_pl)/(floor((T-a))/21)

#LINEAR SHRINKAGE
wolf_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(wolfW[t-a,]<0, 1, 0))/N
    wolf_mon_pl <- cbind(wolf_mon_pl, suma)
  }
}
wolfPL <- sum(wolf_mon_pl)/(floor((T-a))/21)

#Non-linear shrinkage
NL_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(sampleNlW[t-a,]<0, 1, 0))/N
    NL_mon_pl <- cbind(NL_mon_pl, suma)
  }
}
NLinearPL <- sum(NL_mon_pl)/(floor((T-a))/21)

#MARKET FACTOR + LINEAR SHRINKAGE IN RESIDUALS
market_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(marketW[t-a,]<0, 1, 0))/N
    market_mon_pl <- cbind(market_mon_pl, suma)
  }
}
marketPL <- sum(market_mon_pl)/(floor((T-a))/21)

#FAMA FRENCH 3 + LINEAR SHRINKAGE IN RESIDUALS
fama3_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(fama3W[t-a,]<0, 1, 0))/N
    fama3_mon_pl <- cbind(fama3_mon_pl, suma)
  }
}
fama3PL <- sum(fama3_mon_pl)/(floor((T-a))/21)

#FAMA FRENCH 5 + LINEAR SHRINKAGE IN RESIDUALS
fama5_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(fama5W[t-a,]<0, 1, 0))/N
    fama5_mon_pl <- cbind(fama5_mon_pl, suma)
  }
}
fama5PL <- sum(fama5_mon_pl)/(floor((T-a))/21)

#Constant correlation + LINEAR SHRINKAGE IN RESIDUALS
conscor_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(con_wolfW[t-a,]<0, 1, 0))/N
    conscor_mon_pl <- cbind(conscor_mon_pl, suma)
  }
}
conscorPL <- sum(conscor_mon_pl)/(floor((T-a))/21)

#PCA 1 + LINEAR SHRINKAGE IN RESIDUALS
pca1_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(pca1W[t-a,]<0, 1, 0))/N
    pca1_mon_pl <- cbind(pca1_mon_pl, suma)
  }
}
pca1PL <- sum(pca1_mon_pl)/(floor((T-a))/21)

#PCA 3 + LINEAR SHRINKAGE IN RESIDUALS
pca3_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(pca3W[t-a,]<0, 1, 0))/N
    pca3_mon_pl <- cbind(pca3_mon_pl, suma)
  }
}
pca3PL <- sum(pca3_mon_pl)/(floor((T-a))/21)

#PCA 5 + LINEAR SHRINKAGE IN RESIDUALS
pca5_mon_pl <- NULL

for (t in (a+1):T){
  if (t==a+1 | !(t-a-1)%%21){
    suma <- sum(ifelse(pca5W[t-a,]<0, 1, 0))/N
    pca5_mon_pl <- cbind(pca5_mon_pl, suma)
  }
}
pca5PL <- sum(pca5_mon_pl)/(floor((T-a))/21)



PL <- c(demiguelPL, samplePL, wolfPL, NLinearPL, marketPL, fama3PL, fama5PL, conscorPL, pca1PL, pca3PL, pca5PL)
```

```{r}
table1 <- cbind(AV, SD, IR)
rownames(table1) <- c("Equal Weight", "Sample Covariance", "Linear Shrinkage", "Non-linear shrinkage", "Market factor", "Fama-French (3)", "Fama-French(5)", "Constant Correlation","PCA1", "PCA 3", "PCA 5")
(round(table1, 3))

table2 <- cbind(TO, GL, PL)
rownames(table2) <- c("Equal Weight", "Sample Covariance", "Linear Shrinkage", "Non-linear shrinkage", "Market factor", "Fama-French (3)", "Fama-French(5)", "Constant Correlation","PCA1", "PCA 3", "PCA 5")
(round(table2, 2))
```
